---
title: "P8105 Homework 6 [td2704]"
output: github_document
---


```{r, echo=FALSE, message=FALSE}

library(tidyverse)
library(dplyr)

```

## Problem 1

```{r}

birthweight_df = read_csv('data/birthweight.csv')

birthweight_data = 
  birthweight_df %>%
  janitor::clean_names()  %>%
  mutate(
    babysex = as.factor(babysex),
    babysex = fct_recode(babysex, "male" = "1", "female" = "2"),
    frace = as.factor(frace),
    frace = fct_recode(frace, "white" = "1", "black" = "2", 
                       "asian" = "3", "puerto_rican" = "4",
                       "other" = "8"), # "unknown" = "9"
    mrace = as.factor(mrace),
    mrace = fct_recode(mrace, "white" = "1", "black" = "2", 
                       "asian" = "3", "puerto_rican" = "4"), # "other" = "8"
    # malform = as.factor(malform),
    # malform = fct_recode(malform, "absent" = "0", "present" = "1")
    malform = as.logical(malform)
  )

birthweight_data

```

Above, birthweight_data is a cleaned version of the original data with columns babysex, frace,  and mrace turned into factors and column malform turned into a logical ("absent" = "0"/FALSE, "present" = "1"/TRUE).


```{r}

colnames(birthweight_data)[colSums(is.na(birthweight_data)) > 0] 

```

This line above shows that none of the columns have missing values.

Propose a regression model for birthweight. 

My hypothesized structure for the factors that underly birthweight: \
babysex - difference in average birthweight by sex, \
blength - larger length = larger birthweight, \
fincome - higher income = healthier food = larger birthweight, \
malform - weight-related malformations, \
wtgain - mother's weight gain could be correlated, \
gaweeks - gestational age could be correlated

So my linear model will take into account these factors as individual variables to see which variables have the highest correlation.

```{r}
fit = lm(bwt ~ babysex + blength + fincome + malform + wtgain + gaweeks, data = birthweight_data)

fit %>% 
  broom::tidy()

```

Show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

```{r}

birthweight_data %>%
  modelr::add_predictions(fit) %>%
  modelr::add_residuals(fit) %>%
  ggplot(aes(x = pred, y = resid)) + geom_point() +
  ggtitle("Model residuals vs fitted values of birthweight")

```

Let's compare the above model to the following: \
(1) One using length at birth and gestational age as predictors (main effects only). \
(2) One using head circumference, length, sex, and all interactions (including the three-way interaction) between these.

Model (1) One using length at birth and gestational age as predictors (main effects only):

```{r}



```



